### 2020-04-12 Summary.Rmd
Update readme and summary

---
title: "Summary"
author: I. Bartomeus et al. 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Iberian Bee Database.

```{r include=FALSE}
data <- read.csv("data/data_clean.csv")
colnames(data)
```

The Iberian Bee Database contains `r nrow(data)` entries of bee occurrences over the peninsula and nearby islands. We have at least one record of `r length(unique(data$Genus_species))` species out of the 1100? species known to occurr in the area. Here we present a summary of the data base:

## Where are the records?

We have records all over the peninsula, but higer concentrations are marked in red.
```{r echo=FALSE, message=FALSE, warning=FALSE}
#map
library(rgdal)
library(ggplot2)

ESPTmap <- map_data("world", region = c("Portugal", "Spain"))

plotmap <- ggplot(data = data, aes(x = Longitude, y = Latitude)) +  # plot the data
           geom_polygon(data = ESPTmap, aes(x = long, y = lat, group = group),            # plot Spain
                        color = "black", fill = "white") + coord_fixed() +    # coord_fixed() ensures that one unit on the x-axis is the same length as one unit on the y-axis
           geom_point(color = "dodgerblue4", size = 2, shape = ".") +                   # graphical parameters for points
           stat_density2d(aes(x = Longitude,                           # create the density layer based on where the points are
                              y = Latitude,  fill = ..level.., alpha = ..level..),   # colour and transparency depend on density
           geom = "polygon", colour = "grey95",size=0.3) +            # graphical parameters for the density layer
           scale_fill_gradient(low = "yellow", high = "red") +                 # set colour palette for density layer
           scale_alpha(range = c(.25, .5), guide = FALSE) +                    # set transparency for the density layer 
           theme(axis.title.x=element_blank(), axis.text.x=element_blank(),  # don't display x and y axes labels, titles and tickmarks 
           axis.ticks.x=element_blank(),axis.title.y=element_blank(),   
           axis.text.y=element_blank(), axis.ticks.y=element_blank(),
           #text=element_text(size=18),legend.position = c(.9, .15),       # size of text and position of the legend
           panel.grid.major = element_blank(),                            # eliminates grid lines from background
           panel.background = element_blank())                            # set white background
# now plot, it takes a while!
plotmap
```

## Which are the most reported species?

Here we show the top-20 incidence (number of times apearing) per species excluding _Apis mellifera_

```{r echo=FALSE}
#barplot of top 10 species
data2 <- subset(data, Genus_species != "Apis mellifera")
default <- par()$mar
par(mar = c(10.1, 4.1, 4.1, 2.1))
barplot(head(sort(table(data2$Genus_species), decreasing = TRUE),20), las = 2, cex.names = 0.7)
par(mar = default)
```

## When are we recording more bees?

```{r echo=FALSE, warning=FALSE}
hist(data2$Year, main = "Most records are recorded since year 2000", 
     las = 1, xlab = "Year")
#Fix when month is fixed!
hist(as.numeric(as.character(data2$Month)), main = "Most records are recorded in May",
     las = 1, xlab = "Month")
```

## Who make this posible?

This is a large collaborative effort lead by I. Bartomeus, L. Carvalheiro and T. Woods, but it would not be possible without the contribution of thousends of researchers. Those are the main contributors (this can be prettier and more data is in the pipeline)

```{r echo=FALSE, message=FALSE, warning=FALSE}
#table of
#Name	
#Contact	
#Number of records as main dataholder	
#Number of records as co-dataholder
#email?? Maually :(
default <- par()$mar
par(mar = c(13.1, 4.1, 4.1, 2.1))
barplot(head(sort(table(data2$Authors.to.give.credit), decreasing = TRUE),20), las = 2, cex.names = 0.7)
par(mar = default)
#Can be done prettier.
#Ideally, we can merge all authors, split and count individually.
```