### 2022-02-28 Manuscript/Summary/Summary.Rmd
Update on summary

---
#title: 
#author: I. Bartomeus, T. Wood, L. Carvalheiro, C. Molina, et al. - 
#date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
#output: github_document
editor_options: 
  chunk_output_type: console
---

# Summary of the IberianBees database

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(data.table)
data <- read.table("../../Data/iberian_bees.csv.gz",  header=T, quote="\"", sep=",")

master <- read.csv("../../Data/Processing_iberian_bees_raw/Iberian_species_masterlist.csv", stringsAsFactors = FALSE)
master$accepted_name <- paste(master$Genus, master$Species, sep = " ")
colnames(data)
colnames(master)
data[which(!data$accepted_name %in% master$accepted_name),] #OK
master[which(!master$accepted_name %in% data$accepted_name),] 
#we are missing ~100 species, mostly parasites!
length(unique(data$accepted_name))
nrow(master)
```

The Iberian Bee Database contains **`r nrow(data)` entries** of bee occurrences over the peninsula and nearby islands. We have at least one record of **`r length(unique(data$accepted_name))` species** out of the **`r nrow(master)` species** known to occur in the Iberian Peninsula.

We retained only valid, updated species names. All records have accompaining metadata on where and when it was collected, and on who collected, identified and curated the data.

Here we present a summary of the database:

## Where are the records? (Logarithmic scale of the number of occurrences)

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#map
library(rgdal)
library(ggplot2)

ESPTmap <- map_data("world", region = c("Portugal", "Spain"))

plotmap <- ggplot(data = data, aes(x = Longitude, y = Latitude)) +  # plot the data
           geom_polygon(data = ESPTmap, aes(x = long, y = lat, group = group),            # plot Spain
                        color = "black", fill = "white") + coord_fixed() +    # coord_fixed() ensures that one unit on the x-axis is the same length as one unit on the y-axis
           geom_point(color = "dodgerblue4", size = 2, shape = ".") +                   # graphical parameters for points
           stat_density2d(aes(x = jitter(Longitude),                           # create the density layer based on where the points are
                              y = jitter(Latitude),  fill = ..level.., alpha = ..level..),   # colour and transparency depend on density
           geom = "polygon", colour = "grey95",size=0.3) +            # graphical parameters for the density layer
           scale_fill_gradient(low = "yellow", high = "red") +                 # set colour palette for density layer
           scale_alpha(range = c(.25, .5), guide = FALSE) +                    # set transparency for the density layer 
           theme(axis.title.x=element_blank(), axis.text.x=element_blank(),  # don't display x and y axes labels, titles and tickmarks 
           axis.ticks.x=element_blank(),axis.title.y=element_blank(),   
           axis.text.y=element_blank(), axis.ticks.y=element_blank(),
           #text=element_text(size=18),legend.position = c(.9, .15),       # size of text and position of the legend
           panel.grid.major = element_blank(),                            # eliminates grid lines from background
           panel.background = element_blank())                            # set white background
# now plot, it takes a while!
plotmap
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width="70%"}
#code courtesy of Paco Rodriguez.
library(sf)
library(raster)
library(rasterVis)
library(tidyverse)
crs.geo <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
crs.laea <- "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"

library(data.table)
data <- read.table("../../Data/iberian_bees.csv.gz",  header=T, quote="\"", sep=",")

#Prepare data
data$Longitude <- as.numeric(data$Longitude)
data$Latitude <- as.numeric(data$Latitude)
data <- data %>% filter(!is.na(Latitude) & !is.na(Longitude))
#Keep points on balearic islands
data_1 <- data %>% filter(Province=="Balears, Illes")
data_1$Country_1 <- NA

#Filter points in Spain and Portugal
data$Country_1 <- NA
data$Country_1 <- maps::map.where(x = data$Longitude, y = data$Latitude)
data <- data %>% filter(Country_1=="Spain" | Country_1=="Portugal")
#Rbind points that are in Spain with the ones of Balearic islands
dat <- rbind(data, data_1)

datos.sf <- st_as_sf(dat,
                     coords = c("Longitude", "Latitude"),
                     crs = crs.geo)

datos.laea <- st_transform(datos.sf,
                           crs = crs.laea)

ras <- raster(datos.laea, resolution = 10000)

ras.counts <- rasterize(datos.laea[, "accepted_name"], ras,
               field = "accepted_name", fun = "count")
#str(datos.laea)
datos.laea$total <- datos.laea$Male + datos.laea$Female + datos.laea$Worker + datos.laea$Not.specified 
ras.counts2 <- rasterize(datos.laea[, "total"], ras,
               field = "total", fun = "sum")
ras2 <- raster(datos.laea, resolution = 30000)
ras.richness <- rasterize(datos.laea[, "accepted_name"], ras2,
               field = "accepted_name", fun = function(x, ...)
                 {length(unique(na.omit(x)))} )
library(vegan)
ras.chao <- rasterize(datos.laea[, "accepted_name"], ras2,
               field = "accepted_name", fun = function(x, ...)
                 {estimateR(t(as.matrix(x = table(x), nrow = 1, ncol = x)))[2,1]} ) #2 chao, 4 ACE #TO DO THIS WELL, we need to recover abundances.

#levelplot(log(ras.counts+1), margin = FALSE, scales = list(draw = FALSE),
 #         par.settings = viridisTheme())
levelplot(log(ras.counts2+1), margin = FALSE, scales = list(draw = FALSE),
          par.settings = viridisTheme())
#levelplot(log(ras.richness+1), margin = FALSE, scales = list(draw = FALSE),
 #         par.settings = viridisTheme())
#levelplot(ras.chao, margin = FALSE, scales = list(draw = FALSE),
 #         par.settings = viridisTheme())

```

## Which are the 20 most reported species? (Excluding _Apis mellifera_)

```{r echo=FALSE, fig.align='center', out.width="70%"}
#Prepare top 20 records
top_20 <- data %>% filter(!accepted_name=="Apis mellifera") %>% group_by(accepted_name) %>%
summarise(no_rows = length(accepted_name)) %>% 
arrange(-no_rows) %>% slice(1:20)
#Fix row position
top_20$accepted_name <- factor(top_20$accepted_name, levels = top_20$accepted_name)
#Plot
ggplot(top_20, aes(x=accepted_name, y=no_rows)) + 
geom_bar(stat = "identity", fill="#287D8EFF") +
theme_classic()+
theme(axis.text.x=element_text(angle=45, hjust=1, size=8, face = "italic")) +
ylab("Number of records") + 
xlab("Species") + 
scale_y_continuous(expand = c(0, 0))

```

## When are we recording more bees?

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Prepare top 20 records
by_year <- data %>% filter(!accepted_name=="Apis mellifera")

p1 <- ggplot(by_year, aes(Year)) +
  geom_histogram(fill="#287D8EFF") + theme_classic() + ylab("Number of records")+
  scale_y_continuous(expand = c(0, 0))

data$month_name <-  month.abb[data$Month]

month_sum <- data %>% group_by(month_name) %>%
summarise(no_rows = length(month_name)) %>%  arrange(-no_rows)

month_sum <- month_sum %>% filter(!is.na(month_name))

month_sum$month_name <- factor(month_sum$month_name, levels =  c("Jan","Feb","Mar",
              "Apr","May","Jun",
              "Jul","Aug","Sep",
              "Oct","Nov","Dec"))

p2 <- ggplot(data=month_sum, aes(x=month_name, y=no_rows)) +
  geom_bar(stat="identity", fill="#287D8EFF") + theme_classic()+
  scale_y_continuous(expand = c(0, 0)) +
  ylab("Number of records") + xlab("Months")

library(patchwork)

p1 / p2

```

## Who make this posible?

This is a large collaborative effort lead by I. Bartomeus, L. Carvalheiro and T. Woods, but it would not be possible without the contribution of thousends of researchers. Full list of contributors will be available soon.

```{r echo=FALSE,eval=FALSE, message=FALSE, warning=FALSE}
#table of
#Name	
#Contact	
#Number of records as main dataholder	
#Number of records as co-dataholder
#email?? Maually :(
#default <- par()$mar
#par(mar = c(13.1, 4.1, 4.1, 2.1))
#barplot(head(sort(table(data2$Authors.to.give.credit), decreasing = TRUE),20), las = 2, cex.names = 0.7)
#par(mar = default)
#Can be done prettier.
#Ideally, we can merge all authors, split and count individually.
library(stringi)
authors <- data$Authors.to.give.credit[which(!is.na(data$Authors.to.give.credit))]
authors[which(authors == "Ornosa, C., Martínez, M. D.")] <- "C. Ornosa, M.D. Martínez"
authors[which(authors == "C. Ornosa & F. J. Ortiz-Sánchez")] <- "C. Ornosa, F. J. Ortiz-Sánchez"
authors[which(authors == "C. Ornosa & M.D. Martínez")] <- "C. Ornosa, M.D. Martínez"
authors[which(authors == "Ortiz-Sanchez, F.J.")] <- "F.J. Ortiz-Sanchez"
authors[which(authors == "Heras, C., Gayubo, S. F.")] <- "C. Heras, S.F. Gayubo"
authors[which(authors == "Ornosa, C.")] <- "C. Ornosa"
authors[which(authors == "Herrera, J.")] <- "Javier Herrera"
authors[which(authors == "Francisco Javier Ortiz-Sánchez, Pedro Coello & Manuel Baena")] <- "Francisco Javier Ortiz-Sánchez, Pedro Coello, Manuel Baena"
authors[which(authors == "González, J. A., Torres, F., Gayubo, S. F.")] <- "J.A. González, , F. Torres, S.F. Gayubo"
authors[which(authors == "Dardón, M. J., Torres, F., Asís, J. D.")] <- "M.J. Dardón, F. Torres, J.D. Asís"
authors[which(authors == "Pérez-Iñigo Mora, C.")] <- "C. Pérez-Iñigo Mora"
authors[which(authors == "Pablo Vargas, Concepción Ornosa, Francisco Javier Ortiz-Sánchez & Juan Arroyo")] <- "Pablo Vargas, Concepción Ornosa, Francisco Javier Ortiz-Sánchez, Juan Arroyo"
authors[which(authors == "C. Heras y S.F. Gayubo")] <- "C. Heras, S.F. Gayubo"
authors[which(authors == "J. L. BLANCO-PASTOR, C. ORNOSA, D. ROMERO, I. M. LIBERAL, J. M. GÓMEZ & P. VARGAS")] <- "J. L. Blanco-Pastor, C. Ornosa, D. Romero, I.M. Liberal, J.M. Gómez, P. Vargas"
authors[which(authors == "Concepción Ornosa, Félix Torres & Pilar de la Rúa")] <- "Concepción Ornosa, Félix Torres, Pilar de la Rúa"
authors[which(authors == "Castro, L.")] <- "L. Castro"
authors[which(authors == "Ortiz-Sanchez, F. J., Ornosa, C.")] <- "F.J. Ortiz-Sanchez, C. Ornosa"
authors[which(authors == "Banaszak, J., Ortiz-Sanchez, F. J.")] <- "J. Banaszak, F.J. Ortiz-Sanchez"
authors[which(authors == "C. Ornosa, F. Torres & F. J. Ortiz-Sánchez")] <- "C. Ornosa, F. Torres, F. J. Ortiz-Sánchez"

author_pool <- stri_paste(authors, collapse = ",")
author_list <- unlist(strsplit(author_pool, ","))
author_list <- trimws(author_list)
#sort(unique(author_list))[-1] #79 contributors!
#write.csv(sort(unique(author_list))[-1], "data/authors.csv")

auth <- c(
"Alejandro Núñez"
,"Amparo Lázaro"
,"Ana Montero-Castaño"
,"Anna Traveset"
,"Blanca Arroyo-Correa"
,"C. Heras"
,"Curro Molina"
,"C. O’Connor"
,"Concepción Ornosa"
,"C. Pérez-Iñigo Mora"
,"Carlos Lara-Romero"
,"T. Wood"
,"Concepción Ornosa"
,"Concepción Ornosa"
,"D. Paredes"
,"D. Romero"
,"David Alomar"
,"F. Javier Ortiz-Sánchez"
,"F. Javier Ortiz-Sánchez"
,"F. López-Núñez"
,"F. Perfectti"
,"Félix Torres"
,"F. Javier Ortiz-Sánchez"
,"Félix Torres"
,"F. Javier Ortiz-Sánchez"
,"H. Gaspar"
,"Ignasi Bartomeus"
,"Isabel María Liberal"
,"Isabel María Liberal"
,"J. Ascher"
,"J. Banaszak"
,"J. Bosch"
,"J. Dusmet"
,"J. Gonzalez-Varo"
,"J. L. Blanco-Pastor"
,"J. Loureiro"
,"J. R. Obeso"
,"J.A. González"
,"J. Bosch"
,"J.D. Alche"
,"J.D. Asís"
,"J.M. Gómez"
,"Javier Herrera"
,"Javier Morente-López"
,"Joana Costa"
,"José Iriondo"
,"Jose Luis Blanco-Pastor"
,"Juan Arroyo"
,"L. Castro"
,"Luis Navarro"
,"M. Campos"
,"M. Miñarro"
,"M. Porcel"
,"M.D. Martínez"
,"M.J. Dardón"
,"M.L. Fernández"
,"Montserrat Vilà"
,"Manuel Baena"
,"Marina Alonso Solano"
,"Mario Fernández-Mazuecos"
,"Carlos Martínez-Núñez"
,"Miguel A. González-Estévez"
,"Montserrat Vilà"
,"Pablo Vargas"
,"Pablo Vargas"
,"Pedro Coello"
,"Pilar de la Rúa"
,"R. Alcalá"
,"R. Carvalho"
,"R. Heleno"
,"P.J. Rey"
,"S. Castro"
,"S. Castro"
,"S.F. Gayubo"
,"Victoria Ferrero"
,"Volker Haeseler")

unique(auth)
```

Contact nacho.bartomeus@gmail.com if you want to contribute.