### 2022-03-13 Manuscript/Summary/Summary.Rmd
Clean readme, summary and ms

---
#title: "| Base de datos de abejas Ibéricas \n| Iberian Bees database\n"
#date: "`r format(Sys.time(), '%d %B, %Y')`"
output: github_document
editor_options: 
  chunk_output_type: console
---

# Summary of the IberianBees database

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(data.table)
data <- read.table("../../Data/iberian_bees.csv.gz", header = T, quote = "\"", sep = ",")

master <- read.csv("../../Data/Processing_iberian_bees_raw/Iberian_species_masterlist.csv", stringsAsFactors = FALSE)
master$accepted_name <- paste(master$Genus, master$Species, sep = " ")
colnames(data)
colnames(master)
data[which(!data$accepted_name %in% master$accepted_name), ] # OK
master[which(!master$accepted_name %in% data$accepted_name), ]
# we are missing ~100 species, mostly parasites!
length(unique(data$accepted_name))
nrow(master)
```

The Iberian Bee Database contains **`r nrow(data)` entries** of bee occurrences over the peninsula and nearby islands. We have at least one record of **`r length(unique(data$accepted_name))` species** out of the **`r nrow(master)` species** known to occur in the Iberian Peninsula.

We retained only valid, updated species names. All records have accompaining metadata on where and when it was collected, and on who collected, identified and curated the data.

Here we present a summary of the database:

## Where are the records?

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# map
library(rgdal)
library(ggplot2)

ESPTmap <- map_data("world", region = c("Portugal", "Spain"))

plotmap <- ggplot(data = data, aes(x = Longitude, y = Latitude)) + # plot the data
  geom_polygon(
    data = ESPTmap, aes(x = long, y = lat, group = group), # plot Spain
    color = "black", fill = "white"
  ) +
  coord_fixed() + # coord_fixed() ensures that one unit on the x-axis is the same length as one unit on the y-axis
  geom_point(color = "dodgerblue4", size = 2, shape = ".") + # graphical parameters for points
  stat_density2d(aes(
    x = jitter(Longitude), # create the density layer based on where the points are
    y = jitter(Latitude), fill = ..level.., alpha = ..level..
  ), # colour and transparency depend on density
  geom = "polygon", colour = "grey95", size = 0.3
  ) + # graphical parameters for the density layer
  scale_fill_gradient(low = "yellow", high = "red") + # set colour palette for density layer
  scale_alpha(range = c(.25, .5), guide = FALSE) + # set transparency for the density layer
  theme(
    axis.title.x = element_blank(), axis.text.x = element_blank(), # don't display x and y axes labels, titles and tickmarks
    axis.ticks.x = element_blank(), axis.title.y = element_blank(),
    axis.text.y = element_blank(), axis.ticks.y = element_blank(),
    # text=element_text(size=18),legend.position = c(.9, .15),       # size of text and position of the legend
    panel.grid.major = element_blank(), # eliminates grid lines from background
    panel.background = element_blank()
  ) # set white background
# now plot, it takes a while!
plotmap
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width="70%"}
# code courtesy of Paco Rodriguez.
library(sf)
library(raster)
library(rasterVis)
library(tidyverse)
crs.geo <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
crs.laea <- "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"

library(data.table)
data <- read.table("../../Data/iberian_bees.csv.gz", header = T, quote = "\"", sep = ",")

# Prepare data
data$Longitude <- as.numeric(data$Longitude)
data$Latitude <- as.numeric(data$Latitude)
data <- data %>% filter(!is.na(Latitude) & !is.na(Longitude))
# Keep points on balearic islands
data_1 <- data %>% filter(Province == "Balears, Illes")
data_1$Country_1 <- NA

# Filter points in Spain and Portugal
data$Country_1 <- NA
data$Country_1 <- maps::map.where(x = data$Longitude, y = data$Latitude)
data <- data %>% filter(Country_1 == "Spain" | Country_1 == "Portugal")
# Rbind points that are in Spain with the ones of Balearic islands
dat <- rbind(data, data_1)

datos.sf <- st_as_sf(dat,
  coords = c("Longitude", "Latitude"),
  crs = crs.geo
)

datos.laea <- st_transform(datos.sf,
  crs = crs.laea
)

ras <- raster(datos.laea, resolution = 10000)

ras.counts <- rasterize(datos.laea[, "accepted_name"], ras,
  field = "accepted_name", fun = "count"
)
# str(datos.laea)
datos.laea$total <- datos.laea$Male + datos.laea$Female + datos.laea$Worker + datos.laea$Not.specified
ras.counts2 <- rasterize(datos.laea[, "total"], ras,
  field = "total", fun = "sum"
)
ras2 <- raster(datos.laea, resolution = 30000)
ras.richness <- rasterize(datos.laea[, "accepted_name"], ras2,
  field = "accepted_name", fun = function(x, ...) {
    length(unique(na.omit(x)))
  }
)
library(vegan)
ras.chao <- rasterize(datos.laea[, "accepted_name"], ras2,
  field = "accepted_name", fun = function(x, ...) {
    estimateR(t(as.matrix(x = table(x), nrow = 1, ncol = x)))[2, 1]
  }
) # 2 chao, 4 ACE #TO DO THIS WELL, we need to recover abundances.

# levelplot(log(ras.counts+1), margin = FALSE, scales = list(draw = FALSE),
#         par.settings = viridisTheme())
levelplot(log(ras.counts2 + 1),
  margin = FALSE, scales = list(draw = FALSE),
  par.settings = viridisTheme()
)
# levelplot(log(ras.richness+1), margin = FALSE, scales = list(draw = FALSE),
#         par.settings = viridisTheme())
# levelplot(ras.chao, margin = FALSE, scales = list(draw = FALSE),
#         par.settings = viridisTheme())
```

## Which are the 20 most reported species?

```{r echo=FALSE, fig.align='center', out.width="70%"}
# Prepare top 20 records
top_20 <- data %>%
  filter(!accepted_name == "Apis mellifera") %>%
  group_by(accepted_name) %>%
  summarise(no_rows = length(accepted_name)) %>%
  arrange(-no_rows) %>%
  slice(1:20)
# Fix row position
top_20$accepted_name <- factor(top_20$accepted_name, levels = top_20$accepted_name)
# Plot
ggplot(top_20, aes(x = accepted_name, y = no_rows)) +
  geom_bar(stat = "identity", fill = "#287D8EFF") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8, face = "italic")) +
  ylab("Number of records") +
  xlab("Species") +
  scale_y_continuous(expand = c(0, 0))
```

## When are we recording more bees?

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Prepare top 20 records
by_year <- data %>% filter(!accepted_name == "Apis mellifera")

p1 <- ggplot(by_year, aes(Year)) +
  geom_histogram(fill = "#287D8EFF") +
  theme_classic() +
  ylab("Number of records") +
  scale_y_continuous(expand = c(0, 0))

data$month_name <- month.abb[data$Month]

month_sum <- data %>%
  group_by(month_name) %>%
  summarise(no_rows = length(month_name)) %>%
  arrange(-no_rows)

month_sum <- month_sum %>% filter(!is.na(month_name))

month_sum$month_name <- factor(month_sum$month_name, levels = c(
  "Jan", "Feb", "Mar",
  "Apr", "May", "Jun",
  "Jul", "Aug", "Sep",
  "Oct", "Nov", "Dec"
))

p2 <- ggplot(data = month_sum, aes(x = month_name, y = no_rows)) +
  geom_bar(stat = "identity", fill = "#287D8EFF") +
  theme_classic() +
  scale_y_continuous(expand = c(0, 0)) +
  ylab("Number of records") +
  xlab("Months")

library(patchwork)

p1 / p2
```

## Who make this posible?

This is a large collaborative effort lead by I. Bartomeus, L. Carvalheiro and T. Woods, but it would not be possible without the contribution of thousends of researchers.

Contact nacho.bartomeus@gmail.com if you want to contribute.