### 2022-07-01 Manuscript/Figures/Figure_1.Rmd
Update on manuscript figures

---
title: "\\textbf{Figure 1}"
output:
  pdf_document: default
  html_document: default
header-includes:
 - \usepackage{caption}
 - \captionsetup[figure]{labelformat=empty}
 - \captionsetup[table]{labelformat=empty}
---

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',out.width="115%"}
library(sf)
library(raster)
library(rasterVis)
library(tidyverse)
library(data.table)
data <- read.table("../../Data/iberian_bees.csv.gz", header = T, quote = "\"", sep = ",")

# Prepare data
data$Longitude <- as.numeric(data$Longitude)
data$Latitude <- as.numeric(data$Latitude)
data <- data %>% filter(!is.na(Latitude) & !is.na(Longitude))
# Keep points on balearic islands
data_1 <- data %>% filter(Province == "Balears, Illes")
data_1$Country_1 <- NA

# Filter points in Spain and Portugal
data$Country_1 <- NA
data$Country_1 <- maps::map.where(x = data$Longitude, y = data$Latitude)
data <- data %>% filter(Country_1 == "Spain" | Country_1 == "Portugal")
# Rbind points that are in Spain with the ones of Balearic islands
dat <- rbind(data, data_1)

# install.packages("ggplot2")
library(ggplot2)
library(viridis)
library(giscoR) #map of europe
library(ggspatial)

#gisco
#gisco_countrycode
eu2016 <- c(gisco_countrycode[gisco_countrycode$eu, ]$CNTR_CODE)

#Select data
europe_map <- gisco_get_nuts(
  year = "2016",
  epsg = "3035",
  resolution = "3",
  nuts_level = "2",
  country = eu2016) %>%
  sf::st_transform(crs = 4326)

# Borders
borders <- gisco_get_countries(
  epsg = "3035",
  year = "2016",
  resolution = "3",
  country = eu2016) %>%
  sf::st_transform(crs = 4326)

vars <- c("ES",  "PT")
#Extract counties of interest
euro_map <- filter(europe_map, CNTR_CODE %in% vars)

#Plot map
world <- map_data("world")

library(ggsn)

ibera <- world %>% filter(region == "Spain" | region == "Portugal"| region == "France" |region == "Andorra"|region == "Marroco")

ggplot(data=dat, aes(Longitude, Latitude)) +
geom_map(data = ibera, map = ibera,
aes(long, lat, map_id = region), color = "white", fill = "grey", size = 0.1) +
geom_bin2d(bins=70, aes(fill=log(..count..))) +
theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", 
size = 0.5), panel.background = element_rect(fill = "aliceblue"),
panel.border = element_rect(colour = "black", fill=NA, size=1.5)) +
scale_fill_viridis() +
coord_sf(xlim = c(-9, 4), ylim = c(36, 44)) +
ggsn::north(x.min = -0.7, x.max = 3.3, 
y.min = 35.6, y.max = 37.6, scale = 0.4) +
scalebar(x.min = -2, x.max = 3.6,
y.min = 36.5, y.max = 42.5,
dist = 100, dist_unit = "km",st.size = 3,st.dist=0.04,
transform = TRUE, model = "WGS84", st.color="black") +
theme(axis.title = element_text(size=16), axis.text = element_text(size=12))

```