### 2022-07-01 Manuscript/Figures/Figure_2.Rmd

---
title: "\\textbf{Figure 2}"
output:
  pdf_document: default
  html_document: default
header-includes:
 - \usepackage{caption}
 - \captionsetup[figure]{labelformat=empty}
 - \captionsetup[table]{labelformat=empty}
---
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width="100%"}

library(tidyverse)

data <- read.table("../../Data/iberian_bees.csv.gz", header = T, quote = "\"", sep = ",")

# Prepare top 20 records
top_20 <- data %>%
  filter(!Accepted_name == "Apis mellifera") %>%
  group_by(Accepted_name) %>%
  summarise(no_rows = length(Accepted_name)) %>%
  arrange(-no_rows) %>%
  slice(1:20)
# Fix row position
top_20$Accepted_name <- factor(top_20$Accepted_name, levels = top_20$Accepted_name)
# Plot
ggplot(top_20, aes(x = Accepted_name, y = no_rows)) +
  geom_bar(stat = "identity", fill = "#287D8EFF") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 11, face = "italic"), axis.text.y = element_text(size=12),   axis.title=element_text(size=16), axis.line =element_line(size=1)) +
  ylab("Number of records") +
  xlab("Species") +
  scale_y_continuous(expand = c(0, 0))
```