### 2022-07-01 Manuscript/Figures/Figure_3.Rmd

---
title: "\\textbf{Figure 3}"
output:
  pdf_document: default
  html_document: default
header-includes:
 - \usepackage{caption}
 - \captionsetup[figure]{labelformat=empty}
 - \captionsetup[table]{labelformat=empty}
---

```{r echo=FALSE, warning=FALSE, message=FALSE,fig.width=7,fig.height=6, fig.align='center',message=FALSE, warning=FALSE, cache=FALSE, out.width="100%"}

library(tidyverse)
data <- read.table("../../Data/iberian_bees.csv.gz", header = T, quote = "\"", sep = ",")

# Prepare top 20 records
by_year <- data %>% filter(!Accepted_name == "Apis mellifera")

p1 <- ggplot(by_year, aes(Year)) +
  geom_histogram(fill = "#287D8EFF") +
  theme_classic() +
  ylab("Number of records") +
  scale_y_continuous(expand = c(0, 0)) + ggtitle("A") + theme(axis.title=element_text(size=16), axis.text =element_text(size=12), axis.line =element_line(size=1),
plot.title = element_text(size=18, face = "bold"))

data$month_name <- month.abb[data$Month]

month_sum <- data %>%
  group_by(month_name) %>%
  summarise(no_rows = length(month_name)) %>%
  arrange(-no_rows)

month_sum <- month_sum %>% filter(!is.na(month_name))

month_sum$month_name <- factor(month_sum$month_name, levels = c(
  "Jan", "Feb", "Mar",
  "Apr", "May", "Jun",
  "Jul", "Aug", "Sep",
  "Oct", "Nov", "Dec"
))

p2 <- ggplot(data = month_sum, aes(x = month_name, y = no_rows)) +
  geom_bar(stat = "identity", fill = "#287D8EFF") +
  theme_classic() +
  scale_y_continuous(expand = c(0, 0)) +
  ylab("Number of records") +
  xlab("Months") + ggtitle("B") + theme(axis.title=element_text(size=16), axis.text =element_text(size=12), axis.line =element_line(size=1),
plot.title = element_text(size=18, face = "bold"))

library(patchwork)

p1 / p2
```